Redundant interferometric calibration as a complex optimization problem

### 2018-02-08 paper.tex
second round of corrections

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%
\usepackage{subfigure}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{graphicx}
%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}
\usepackage{mathrsfs}
\usepackage{url}

@@

The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we presnt our conclusions in Section~\ref{sec:conclusions}.

@@

\begin{table}
\centering
\caption{Notation and frequently used symbols.}
\begin{tabular}{|l l|} 
\hline
$x$, $\mathbf{x}$ and $\mathbf{X}$ & scalar, vector and matrix\\
$\mathbb{N}$ & set of natural numbers\\
$\mathbb{C}$ & set of complex numbers\\
$\bI$ & Identity matrix\\
$\alpha_{pq}$, $\phi_{pq}$, $\zeta_{pq}$, $\xi_{pq}$ and & indexing functions\\
$\psi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}$ &\\
$\conj{(*)}$ & conjugation of operand\\
$(*)^{2}$ & square operand\\
$\|*\|_F$ & Frobenius norm\\
$(*)^{-1}$ & Matrix inversion\\
$\odot$ & Hadamard product\\
&(element-wise product)\\
$(*)^H$ & Hermitian transpose\\
$(*)^T$ & Matrix transpose\\
$<*>_x$ & averaging in x\\
$\frac{\partial}{\partial z}$ & Wirtinger derivative\\
$\widetilde{x}$ & a new quantity derived from $x$\\
$\widehat{x}$ & estimated quantity\\
$\breve{\mathbf{x}}$ & augmented vector $[\mathbf{x}^T,\conj{\mathbf{x}}^T]^T$\\
$i$ & $\sqrt{-1}$\\
$N$,$B$,$L$ and $P$ & number of antennas, baselines,\\
&redundant groups and parameters\\
$\bd$, $\bv$, $\bg$, $\by$, $\br$ and $\bn$& data, predicted, antenna,\\
&true visibility, residual\\
&and noise vector.\\ 
$\bz$ & $\bz = [\bg^T,\by^T]^T$\\
$\Lambda$ & objective function\\
$\lambda$ and $\rho$ & damping factor and $\rho = \frac{1}{1+\lambda}$\\
$\bJ$ & Jacobian matrix\\
$\bH$ & $\bJ^H\bJ$ (Hessian matrix)\\
$\bmH$ & modified Hessian matrix\\
$[\mathbf{X}]_{ij}$& element $ij$ of matrix $\mathbf{X}$\\
$x_{ij}$ & composite antenna index\\
$x_i$ & antenna or redundant group index\\
$x^k$ & iteration number\\
$\bM$ & preconditioner matrix\\
$m$ & number of non-zero entries\\
&in a matrix\\
$\kappa$ & spectral condition number\\
&of a matrix\\
$\gamma$ & sparsity factor of a matrix\\
$\Delta x$ & parameter update\\
\hline
\end{tabular}
\label{tab:glossary}
\end{table}

@@

In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$ (i.e. an antenna gain). The observed visibility $d_{pq}$ is therefore given by \citep{ME1,ME2,RRIME1}
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q} \, y_{pq} + n_{pq},
\end{equation}
where $\conj{(*)}$ indicates complex conjugation and $n_{pq}$ is the thermal noise component. 
The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
standard deviation $\sigma$:
\begin{equation}
\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}},
\end{equation}
where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.

Considering the number of visibilities $B$ (i.e. baselines) measured by an array of $N$ elements $B = \frac{N^2-N}{2}$, equation~\ref{eq:vis_definition} can be expressed in the following vector form:
\begin{equation}
\label{eq:vis_linear_definition}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_linear_definitions}
\end{align}
and 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
The function $\alpha_{pq}$ therefore maps composite antenna indexes to unique single indexes, i.e:
\begin{equation}
\{\alpha_{12},\alpha_{13},\cdots,\alpha_{N-1N}\} = \{1,2,\cdots,B\} 
\end{equation}
The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$). It is important to point out here that 
all of the mathematical definitions in this paper assumes that we are only considering composite indices belonging to the set $\{rs|r<s\}$. Fore the sake of mathematical 
completeness, however, $\alpha_{pq}$ is defined for composite indices which do not belong to this set.

Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
\begin{equation}
\label{eq:least_squares}
\min_{\bg} \Lambda(\bg) = \min_{\bg} \|\br\|_F^2 = \min_{\bg} \|\bd - \bv(\bg)\|_F^2, 
\end{equation}
where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|_F$ denotes the Frobenius norm. In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.

@@

In this paper, we use single subscript indices (e.g. $i$) to refer to a specific antenna (or as will become apparent later a specific redundant group) and composite subscript indices (e.g. $pq$) to refer to 
a specific baseline. If these indices appear in the definition of matrix elements then their allowed ranges are determined by the dimension of the matrix that 
is being defined (see Table~\ref{tab:matrix_dimensions_main}). Furthermore, the identity matrix is denoted by $\bI$. Moreover, the Hadamard\footnote{element-wise product} product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$ respectively \citep{Liu2008}. Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
We will refer to $\bJ^H\bJ$ as the Hessian\footnote{a square matrix composed of second-order partial derivatives} matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper \citep{Madsen:LM}.

@@

\begin{figure}
\includegraphics[width=0.47\textwidth]{./flow_diagram.pdf}
\caption{A flow chart representing the procedure one would follow to partially construct the Jacobian matrix (i.e. $\bmJ$) associated with Wirtinger redundant calibration. \label{fig:flow_chart}}
\end{figure}
%\cite{Liu2008}\cite{Madsen:LM}

@@

Note that we employ the same subscript indexing notation that we used in Section~\ref{sec:sky_wirtinger} in equation~\ref{eq:j1} and equation~\ref{eq:j2}.
To further aid the reader in understanding this indexing notation we refer him/her to Figure~\ref{fig:flow_chart} which depicts a flow diagram of the matrix construction procedure 
with which $\bmJ_1$ can be constructed. The range of values the indexes in equation~\ref{eq:j1} and equation~\ref{eq:j2} can attain should also 
be clear to the reader after having inspected Figure~\ref{fig:flow_chart} and Table~\ref{tab:matrix_dimensions_main}.

@@

We also did not explicitly define our noise in terms of the integration time, channel bandwidth and $T_{\textrm{sys}}$, we instead
follow the same approach described in \citep{Liu2010,Marthi2014} and make use of the definition of SNR to introduce noise into our visibilities. 
We use the following definition of 
\citep[SNR, ][]{Liu2010,Marthi2014}:  
\begin{equation}
\label{eq:SNR}
\textrm{SNR} = 10\log \left (\frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}} \right ), 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. It should be pointed out, however, that by producing visibilities with different SNR values you are 
effectively either changing your integration time or your channel bandwidth or both (assuming a single $T_{\textrm{sys}}$ value for your instrument). The reason we decided to also adopt
equation~\ref{eq:SNR} is that it has become standard practice to present the accuracy of parameter estimation algorithms as a function of SNR.

@@

\section{Conclusion}
\label{sec:conclusions}
In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism 
\citep{Smirnov2015}. We derived the associated complex Jacobian and the GN and LM parameter update steps.
We also showed that the LM parameter update step can be simplified to obtain an ADI type of algorithm \citep[][]{Salvini2014,Marthi2014}.
Our code implementation of this algorithm (redundant \textsc{StEfCal}) is publicly available at 
\url{https://github.com/Trienko/heracommissioning/blob/master/code/stef.py}. 
We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work. Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for easier parallelization.

@@

\section{Future Outlook}
We plan to apply redundant \textsc{StEfCal} to HERA data in the near future. The reason for applying redundant \textsc{StEfCal} to real data is two-fold. First, we wish to validate the theoretical computational
complexity of redundant \textsc{StEfCal} (derived earlier in the paper). Second, we would like to test whether it could be used to calibrate HERA in near-realtime. 
We are also interested in paralyzing our current implementation and to see how much can be gained by doing so. We would also like to 
conduct a perturbation analyses, similar to the one conducted by \cite{Liu2010}, to estimate the error which is introduced into our 
estimated gains and visibilities when the array contains baselines which are not perfectly redundant. We are also interested in quantifying 
the error which is introduced into the estimated gains and visibilities because of differing primary beam patterns between array elements \citep{Noorishad2012}.

@@

% \section{Conjugate Gradient Method}
% 
% \subsection{Normal Equation}
% \label{sec:normal}
% Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
% \begin{equation}
%  \bb = \bB\bx,
% \end{equation}
% where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
% \begin{equation}
% \label{eq:normal_equation}
% \bB^H\bB\bx = \bB^H\bb, 
% \end{equation}
% to estimate the $\bx$ that minimizes
% \begin{equation}
% \label{eq:norm_ls}
% \|\bb-\bB\bx\|_F^2. 
% \end{equation}
% In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
% 
% There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
% \begin{enumerate}
% \item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
% \item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
% \begin{equation}
% \bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
% \end{equation}
% \end{enumerate}
% In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
% 
% Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability: it can be used to solve equation~\ref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
% square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}. In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that equation~\ref{eq:normal_equation}
% can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
% obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
% 
% Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
% is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
% number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
% $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.

@@

\begin{equation}
[\bC]_{ij} = 
\begin{cases}
 \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
[\bD]_{ij} = 
\begin{cases}
 g_i \conj{y}_j  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}

>>>>

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%
\usepackage{subfigure}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{mathrsfs}
\usepackage{url}

@@

The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we present our conclusions in Section~\ref{sec:conclusions}.

@@

\begin{table}
\centering
\caption{Notation and frequently used symbols.}
\begin{tabular}{|l l|} 
\hline
$x$, $\mathbf{x}$ and $\mathbf{X}$ & scalar, vector and matrix\\
$\mathbb{N}$ & set of natural numbers\\
$\mathbb{C}$ & set of complex numbers\\
$\bI$ & Identity matrix\\
$\alpha_{pq}$, $\phi_{pq}$, $\zeta_{pq}$, $\xi_{pq}$ and & indexing functions\\
$\psi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}$ &\\
$\conj{x}$ & conjugation\\
$\conj{\mathbf{x}},\conj{\mathbf{X}}$ & element-wise conjugation\\
$x^{2}$ & square\\
$\|\mathbf{x}\|_F$ & Frobenius norm\\
$\mathbf{X}^{-1}$ & Matrix inversion\\
$\odot$ & Hadamard product\\
&(element-wise product)\\
$\mathbf{X}^H$ & Hermitian transpose\\
$\mathbf{x}^T,\mathbf{X}^T$ & Transpose\\
$<\mathbf{x}>_x$ & averaging over $x$\\
$\frac{\partial}{\partial z}$ & Wirtinger derivative\\
$\widetilde{x}$ & a new quantity derived from $x$\\
$\widehat{x}$ & estimated quantity\\
$\breve{\mathbf{x}}$ & augmented vector $[\mathbf{x}^T,\conj{\mathbf{x}}^T]^T$\\
$i$ & $\sqrt{-1}$\\
$N$,$B$,$L$ and $P$ & number of antennas, baselines,\\
&redundant groups and parameters\\
$\bd$, $\bv$, $\bg$, $\by$, $\br$ and $\bn$& data, predicted, antenna,\\
&true visibility, residual\\
&and noise vector.\\ 
$\bz$ & $\bz = [\bg^T,\by^T]^T$\\
$\Lambda$ & objective function\\
$\lambda$ and $\rho$ & damping factor and $\rho = \frac{1}{1+\lambda}$\\
$\bJ$ & Jacobian matrix\\
$\bH$ & $\bJ^H\bJ$ (Hessian matrix)\\
$\bmH$ & modified Hessian matrix\\
$[\mathbf{X}]_{ij}$& element $ij$ of matrix $\mathbf{X}$\\
$x_{ij}$ & composite antenna index\\
$x_i$ & antenna or redundant group index\\
$x^k$ & iteration number\\
$\bM$ & preconditioner matrix\\
$m$ & number of non-zero entries\\
&in a matrix\\
$\kappa$ & spectral condition number\\
&of a matrix\\
$\gamma$ & sparsity factor of a matrix\\
$\Delta x$ & parameter update\\
\hline
\end{tabular}
\label{tab:glossary}
\end{table}

@@

In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$ (i.e. an antenna gain). The observed visibility $d_{pq}$ is therefore given by \citep{ME1,ME2,RRIME1}
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q} \, y_{pq} + n_{pq},
\end{equation}
where $\conj{\conj{x}}$ indicates complex conjugation and $n_{pq}$ is the thermal noise component. 
The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
standard deviation $\sigma$:
\begin{equation}
\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}},
\end{equation}
where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.

Considering the number of visibilities $B$ (i.e. baselines) measured by an array of $N$ elements $B = \frac{N^2-N}{2}$, equation~\ref{eq:vis_definition} can be expressed in the following vector form:
\begin{equation}
\label{eq:vis_linear_definition}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_linear_definitions}
\end{align}
and 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
The function $\alpha_{pq}$ therefore maps composite antenna indexes to unique single indexes, i.e:
\begin{equation}
\{\alpha_{12},\alpha_{13},\cdots,\alpha_{N-1N}\} = \{1,2,\cdots,B\} 
\end{equation}
The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$). It is important to point out here that 
all of the mathematical definitions in this paper assumes that we are only considering composite indices belonging to the set $\{rs|r<s\}$. For the sake of mathematical 
completeness, however, $\alpha_{pq}$ is defined for composite indices which do not belong to this set. Also note that, while the indexing function $\alpha_{pq}$ does not attain 
the value zero in practice the zero indexing value does play a very important role in some of the definitions used within this paper (see for example Eq.~\ref{eq:D_mat}).

Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
\begin{equation}
\label{eq:least_squares}
\min_{\bg} \Lambda(\bg) = \min_{\bg} \|\br\|_F^2 = \min_{\bg} \|\bd - \bv(\bg)\|_F^2, 
\end{equation}
where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\mathbf{x}\|_F$ denotes the Frobenius norm. In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.

@@

In this paper, we use single subscript indices (e.g. $i$) to refer to a specific antenna (or as will become apparent later a specific redundant group) and composite subscript indices (e.g. $pq$) to refer to 
a specific baseline. If these indices appear in the definition of matrix elements then their allowed ranges are determined by the dimension of the matrix that 
is being defined (see Table~\ref{tab:matrix_dimensions_main}). Furthermore, the identity matrix is denoted by $\bI$. Moreover, the Hadamard\footnote{element-wise product} product and the Hermition transpose are denoted by $\odot$ and $\mathbf{X}^H$ respectively \citep{Liu2008}. Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
We will refer to $\bJ^H\bJ$ as the Hessian\footnote{a square matrix composed of second-order partial derivatives} matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper \citep{Madsen:LM}.

@@

\begin{figure}
\includegraphics[width=0.47\textwidth]{./flow_diagram.pdf}
\caption{A flow chart representing the procedure one would follow to partially construct the Jacobian matrix (i.e. $\bmJ_1$) associated with Wirtinger redundant 
calibration. The red circle represents the start of the flow diagram. The blue circle represents the end of the diagram. Blue diamonds denote loop 
conditionals, while green diamonds denote simple conditionals. The diagram elements following the red arrows just below a loop conditional element all form part 
of the main body of the loop which has the loop conditional of the aforementioned loop conditional element in its definition. The pseudocode associated with this flow diagram is given in Algorithm 1. \label{fig:flow_chart}}
\end{figure}

@@

\begin{algorithm}
\caption{Constructing $\bmJ_1$}\label{euclid}
\begin{algorithmic}[1]
\State $p \gets 1$
\While {$p\leq N$}
\State $q \gets p + 1$
\While {$q\leq N$}
\State $\alpha  \gets (q-p) + (p-1)\left (N-\frac{1}{2} \right )$
\State $i \gets 1$
\While {$i\leq (N + L)$}
\If {$i \leq N$}
\State $\bmJ_1[\alpha,i] \gets \frac{\partial v_{pq}}{\partial g_i}$
\Else
\State $\bmJ_1[\alpha,i] \gets \frac{\partial v_{pq}}{\partial y_{i-N}}$
\EndIf
\State $i\gets i + 1$
\EndWhile
\State $q \gets q + 1$
\EndWhile
\State $p \gets p + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

Note that we employ the same subscript indexing notation that we used in Section~\ref{sec:sky_wirtinger} in equation~\ref{eq:j1} and equation~\ref{eq:j2}.
To further aid the reader in understanding this indexing notation we refer him/her to Figure~\ref{fig:flow_chart} (also see Algorithm 1) which depicts a flow diagram of the matrix construction procedure 
with which $\bmJ_1$ can be constructed. The range of values the indexes in equation~\ref{eq:j1} and equation~\ref{eq:j2} can attain should also 
be clear to the reader after having inspected Figure~\ref{fig:flow_chart} and Table~\ref{tab:matrix_dimensions_main}.

@@

We also did not explicitly define our noise in terms of the integration time, channel bandwidth and $T_{\textrm{sys}}$, we instead
follow the same approach described in \citep{Liu2010,Marthi2014} and make use of the definition of SNR to introduce noise into our visibilities. 
We use the following definition of 
\citep[SNR, ][]{Liu2010,Marthi2014}:  
\begin{equation}
\label{eq:SNR}
\textrm{SNR} = 10\log \left (\frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}} \right ), 
\end{equation}
where $<\mathbf{x}>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. It should be pointed out, however, that by producing visibilities with different SNR values you are 
effectively either changing your integration time or your channel bandwidth or both (assuming a single $T_{\textrm{sys}}$ value for your instrument). The reason we decided to also adopt
equation~\ref{eq:SNR} is that it has become standard practice to present the accuracy of parameter estimation algorithms as a function of SNR.

@@

\section{Conclusion}
\label{sec:conclusions}
In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism 
\citep{Smirnov2015}. We derived the associated complex Jacobian and the GN and LM parameter update steps.
We also showed that the LM parameter update step can be simplified to obtain an ADI type of algorithm \citep[][]{Salvini2014,Marthi2014}.
Our code implementation of this algorithm (redundant \textsc{StEfCal}) is publicly available at 
\url{https://github.com/Trienko/heracommissioning/blob/master/code/stef.py}. 
We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016,dillon2017} which will be the subject of future work. Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for easier parallelization.

@@

\section{Future Outlook}
We plan to apply redundant \textsc{StEfCal} to HERA data in the near future. The reason for applying redundant \textsc{StEfCal} to real data is two-fold. First, we wish to validate the theoretical computational
complexity of redundant \textsc{StEfCal} (derived earlier in the paper). Second, we would like to test whether it could be used to calibrate HERA in near-realtime. 
We are also interested in parallelizing our current implementation and to see how much can be gained by doing so. We would also like to 
conduct a perturbation analyses, similar to the one conducted by \cite{Liu2010}, to estimate the error which is introduced into our 
estimated gains and visibilities when the array contains baselines which are not perfectly redundant. We are also interested in quantifying 
the error which is introduced into the estimated gains and visibilities because of differing primary beam patterns between array elements \citep{Noorishad2012}.

@@

% \section{Conjugate Gradient Method}
% 
% \subsection{Normal Equation}
% \label{sec:normal}
% Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
% \begin{equation}
%  \bb = \bB\bx,
% \end{equation}
% where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
% \begin{equation}
% \label{eq:normal_equation}
% \bB^H\bB\bx = \bB^H\bb, 
% \end{equation}
% to estimate the $\bx$ that minimizes
% \begin{equation}
% \label{eq:norm_ls}
% \|\bb-\bB\bx\|_F^2. 
% \end{equation}
% In this paper, we denote the Hermitian transpose with $(\cdot)^H$ and the Frobenius norm with $\|*\|_F$.
% 
% There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
% \begin{enumerate}
% \item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
% \item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
% \begin{equation}
% \bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
% \end{equation}
% \end{enumerate}
% In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
% 
% Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability: it can be used to solve equation~\ref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
% square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}. In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that equation~\ref{eq:normal_equation}
% can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
% obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
% 
% Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
% is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
% number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
% $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.

@@

\begin{equation}
[\bC]_{ij} = 
\begin{cases}
 \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
\label{eq:D_mat}
[\bD]_{ij} = 
\begin{cases}
 g_i \conj{y}_j  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
